---
title: "Exercise 1"
output:
  pdf_document:
    latex_engine: pdflatex
  html_document:
    df_print: paged
header-includes:
- \usepackage{helvet}
- \usepackage[T1]{fontenc}
---

```{r}
df <- read.csv("C:/Users/csg20/Downloads/ppl analytic/Connections.csv", header = TRUE)
colnames(df) <- gsub("\\.", "_", colnames(df))
colnames(df)
```

```{r}
library(dplyr)
library(tidyverse)
library(tidygraph)
library(ggraph)
library(igraph)
library(readr)
```

```{r}
df <- df %>%
  filter(!((is.na(First_Name) | First_Name == "") & (is.na(Last_Name) | Last_Name == "")))

df <- df %>%
  distinct(First_Name, Last_Name, .keep_all = TRUE)
```

```{r}
count_by_employer <- df %>%
  group_by(Company) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count))

# Display the counts
print(count_by_employer)
```

```{r}
df %>%
  count()
```

```{r}
# Create nodes dataframe with unique individuals
nodes <- df %>%
  mutate(Label = paste(First_Name, substr(Last_Name, 1, 1), sep = " ")) %>%
  select(Label, Company) %>%
  distinct() %>%
  filter(!is.na(Label))
```

```{r}
# Create edges based on shared company affiliation
edges <- nodes %>%
  select(Label, Company) %>%
  distinct() %>%
  inner_join(nodes %>% select(Label, Company) %>% distinct(), by = "Company", suffix = c("_x", "_y")) %>%
  filter(Label_x != Label_y) %>%
  select(from = Label_x, to = Label_y) %>%
  distinct()

# Ensure there are no NA values and that 'from' and 'to' columns are properly matched to node IDs
edges <- edges %>%
  filter(!is.na(from) & !is.na(to)) %>%
  left_join(nodes, by = c("from" = "Label")) %>%
  left_join(nodes, by = c("to" = "Label"))
```

```{r}
tidy_graph <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE)

tidy_graph <- tidy_graph %>% 
  activate(nodes) %>%
  mutate(community = as.factor(group_louvain()))

# Create the graph visualization
g <- ggraph(tidy_graph, layout = 'kk') + 
  geom_edge_link(aes(alpha = stat(index)), show.legend = FALSE) + 
  geom_node_point(aes(colour = community), show.legend = FALSE, size = 1) + 
  theme_graph()

g
```

```{r}
nodes <- df %>%
  mutate(Label = paste(First_Name, substr(Last_Name, 1, 1), sep = " ")) %>%
  select(Label, Company) %>%
  distinct() %>%
  filter(!is.na(Label)) %>%
  mutate(group = case_when(
    Company == "McGill University - Desautels Faculty of Management" ~ "Desautels",
    Company == "McGill University" ~ "McGill",
    TRUE ~ "Other"))


tidy_graph <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE)

tidy_graph <- tidy_graph %>% 
  activate(nodes) %>%
  mutate(community = as.factor(group_louvain()))

# Create the graph visualization
g <- ggraph(tidy_graph, layout = 'fr') + 
  geom_edge_link() + 
  geom_node_point(aes(color = group), size = 1) + 
  theme_graph()+
  scale_color_manual(values = c("Desautels" = "orange", "McGill" = "red", "Other" = "blue")) +
  theme_graph(base_family="sans") +
  labs(color = "Group")

# Print the graph
g
```

The clustering of Desautels and McGill nodes indicates a dense network of connections within university and faculty, which is expected given I am studying at Desautels at McGill. The 'Other' category represents connections from different facets such as past educational institutions, colleagues, friends, or industry professionals not directly associated with McGill or Desautels. The Desautels cluster is small and isolated, indicating little to no direct interaction with nodes from other groups. In contrast, the McGill cluster is closer to the 'Other' nodes, suggesting some level of interaction or connection between the McGill group and the 'Other' group.
